---
title: "Case Study 1: Simulation Recovery"
author: "David Hodgson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case Study 1: Simulation Recovery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# R Code Description

This document describes the steps and components of the provided R code for creating and running serological models using simulated data. The code defines observational models, antibody kinetics models, and runs inference using MCMC methods. 

---

## Step 1: Load Required Libraries
```{r, message=FALSE, warning=FALSE, results='hide'}

library(devtools)
library(tidyr)
library(dplyr)
library(magrittr)
library(serojump)
#devtools::load_all()
library(patchwork)

# Using more than one core might fail on windows, 
if (.Platform$OS.type == "windows") {
  mc.cores <- 1
} else {
  mc.cores <- 2 # use as many as available, preferably 4.
}


```

## Step 2: Read Simulated Data

This data has been generated using `serosim`. The code the generate this data can be found at in the folder `R/serosim` in the package. The data is generated using the `cesCOP` and `cesNoCOP` models.

```{r}


sim_model_cop <- readRDS(file = "cesCOP_inputs.RDS")
sim_res_cop <- readRDS(file = "cesCOP_sim_data_0.5.rds")

#modeli <- readRDS(file = system.file("extdata", "vig_data", "cesCOP", "cesCOP_inputs.RDS", package = "serojump"))
#res <- readRDS(file = system.file("extdata","vig_data", "cesCOP", "cesCOP_sim_data_0.5.RDS", package = "serojump"))

#modeli <- readRDS(file = "cesCOP_inputs.RDS")
#modeli <- readRDS(file = "cesCOP_sin_data_0.5.RDS")

data_titre_model <- sim_res_cop$observed_biomarker_states %>% select(i, t, value) %>% rename(id = i, time = t, titre = value) 
data_titre_model <- data_titre_model %>% mutate(biomarker = "sVNT") %>% as.data.frame %>% rename(sVNT = titre)

# Check the entries are sensible
check_sero_no_single_entries(data_titre_model)
check_sero_timings(data_titre_model)

```

## Step 3: Define Likelihood and Kinetics Functions

```{r}

obsLogLikelihood <- function(titre_val, titre_est, pars) {
    ll <- dnorm(titre_val, titre_est, pars[1], log = TRUE)
}

infSerumKinetics <- function(titre_est, timeSince, pars) {
    a <- pars[1]
    b <- pars[2]
    c <- pars[3]
    if (timeSince < 14) {
        titre_est <- titre_est + log(exp(a) + exp(c)) * (timeSince) / 14;
    } else {
        titre_est <- titre_est + log(exp(a) * exp(-b/10 * (timeSince - 14)) + exp(c));
    }
    titre_est
}

noInfSerumKinetics <- function(titre_est, timeSince, pars) {
    titre_est_log <- titre_est - pars[1] * (timeSince)
    titre_est_log <- max(0, titre_est_log)
    titre_est_log
}

```

## Step 4: Define the Observational and Kinetics Models

```{r}


# Define the biomarkers and exposure types in the model
biomarkers <- c("sVNT")
exposureTypes <- c("none", "inf")
exposureFitted <- "inf"

# Define the observational model
observationalModel <- list(
    names = c("sVNT"),
    model = makeModel(
        addObservationalModel("sVNT", c("sigma"), obsLogLikelihood)
        ), # observational model,
    prior = bind_rows(
            addPrior("sigma", 0.0001, 4, "unif", 0.0001, 4)
    )
)

# Define the antibody kinetics model
abkineticsModel <- list(
    model = makeModel(
            addAbkineticsModel("none", "sVNT", "none",  c("wane"), noInfSerumKinetics),
            addAbkineticsModel("inf", "sVNT", "inf", c("a", "b", "c"), infSerumKinetics)
        ),
    prior = bind_rows(
        addPrior("wane", 0.0, 0.01, "unif", 0.0, 0.01), # observational model
        addPrior("a", -6, 6, "norm",  2, 2), # ab kinetics
        addPrior("b", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        addPrior("c", 0, 4, "unif", 0,  4) # ab kinetics 
    )
)



model_cop <- createSeroJumpModel(
    data_sero = data_titre_model, 
    data_known = NULL,
    biomarkers = biomarkers,
    exposureTypes = exposureTypes,
    exposureFitted = exposureFitted,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel)


```


## Step 4B: Prerun sanity plots

Before running the whole model it is good to check the data and the priors. This can be done using a suit of functions `plotPriors` function.

```{r}

p1 <- plotSero(model_cop)
p2 <- plotPriorPredictive(model_cop)
p3 <- plotPriorInfection(model_cop)
p1 / p2
p3

```

## Step 5: Define the Complete Model



```{r}

sim_model_no_cop <- readRDS(file = "cesNoCOP_inputs.RDS")
sim_res_no_cop <- readRDS(file = "cesNoCOP_sim_data_0.5.rds")

#modeli <- readRDS(file = system.file("extdata", "vig_data", "cesNoCOP", "cesNoCOP_inputs.RDS", package = "serojump"))
#res <- readRDS(file = system.file("extdata", "vig_data", "cesNoCOP", "cesNoCOP_sim_data_0.5.RDS", package = "serojump"))

data_titre_model <- sim_res_no_cop$observed_biomarker_states %>% select(i, t, value) %>% rename(id = i, time = t, titre = value) 
data_titre_model <- data_titre_model %>% mutate(biomarker = "sVNT") %>% as.data.frame %>% rename(sVNT = titre)


model_no_cop <- createSeroJumpModel(
    data_sero = data_titre_model, 
    data_known = NULL,
    biomarkers = biomarkers,
    exposureTypes = exposureTypes,
    exposureFitted = exposureFitted,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel)

```


## Step 5B: Prerun sanity plots

Before running the whole model it is good to check the data and the priors. This can be done using a suit of functions `plotPriors` function.

```{r}

p1 <- plotSero(model_no_cop)
p2 <- plotPriorPredictive(model_no_cop)

p1 / p2

```

```{r, message=FALSE, warning=FALSE, results='hide'}


rj_settings <- list(
        numberChainRuns = 4, 
        iterations = 400000,
        burninPosterior = 200000,
        thin = 1000
    )

save_info_cop <- list(
    file_name = "simulated_data",
    model_name = "cop"
)

save_info_no_cop <- list(
    file_name = "simulated_data",
    model_name = "no_cop"
)

# Run these but take a while
#model_summary_cop <- runSeroJump(model_cop, rj_settings, save_info = save_info_cop)
#model_summary_no_cop <- runSeroJump(model_no_cop, rj_settings, save_info = save_info_no_cop)

model_summary_cop <-  readRDS("cop_model_summary.RDS")
model_summary_no_cop <-  readRDS("no_cop_model_summary.RDS")

plotMCMCDiagnosis(model_summary_cop, save_info = save_info_cop)
# takes ages to run these so comment out!
#plotPostFigsSim(model_summary_cop, sim_model_cop, sim_res_cop, save_info = save_info_cop)

plotMCMCDiagnosis(model_summary_no_cop, save_info = save_info_no_cop)
# takes ages to run these so comment out!
#plotPostFigsSim(model_summary_no_cop, sim_model_no_cop, sim_res_no_cop, save_info = save_info_no_cop)

```