---
title: "Intro 2: Explanation of observational and antibody kinetics models"
author: "David Hodgson"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

This document explains two critical parts of the model defined in the provided R code:

1. The **Observational Model**: This component models the observed data, specifying how the measured values relate to the latent (unobserved) processes.
2. The **Antibody Kinetics Model**: This component models the changes in antibody levels over time, accounting for both natural decay (waning) and infection-induced changes.

---

## 0. The Observational Model

```{r}

library(devtools)
library(tidyverse)
devtools::load_all()
#library(serojump)


```

## 1. The Observational Model

### Definition

The observational model captures the process by which observed data (`PreF` in this case) is generated from the underlying system. It defines the relationship between the true (latent) variables and the observed data.

### Code Breakdown

```{r}

obsLogLikelihood <- function(titre_val, titre_est, pars) {
    ll <- dnorm(titre_val, titre_est, pars[1], log = TRUE)
    ll
}

observationalModel <- list(
    names = c("IgG"),
    model = makeModel(
        addObservationalModel("IgG", c("sigma"), obsLogLikelihood)
    ),
    prior = bind_rows(
        add_par_df("sigma", 0.0001, 4, "unif", 0.0001, 4)
    ) # observational model,
)

```

- **`names = c("IgG")`**: The observed biomarker here is `IgG`, representing some measurement, such as a biomarker (e.g., antibody concentration).

- **`model = makeModel(...)`**: The model is created using the `makeModel` function, which defines how the observation model is constructed.

- **`addObservationalModel("IgG", c("sigma"), obsLogLikelihood)`**: This line adds an observational model for `IgG`, where the parameter `sigma` is used to model the noise or uncertainty in the observations. The log-likelihood function (`obsLogLikelihood`) is used to define how the observations relate to the latent variable.

- **`prior = bind_rows(...)`**: This block defines the prior distribution for the parameter `sigma`, which controls the observation noise.

- **`add_par_df("sigma", 0.0001, 4, "unif", 0.0001, 4)`**: The parameter `sigma` is given a uniform prior distribution between `0.0001` and `4`, indicating that the model expects a certain range of variability in the observed data.

### Summary

The observational model defines how the `IgG` variable is observed with some noise, controlled by `sigma`. The log-likelihood function helps relate the true underlying values to the observed values, while the prior on `sigma` assumes uniform uncertainty.


## 2. The Antibody Kinetics Model

### Definition

The antibody kinetics model describes how antibody levels evolve over time in response to different factors, such as infection or natural waning. This model accounts for both natural decay in antibody levels and changes due to infection.

### Code Breakdown
Define the antibody kinetic functions
```{r}


noInfSerumKinetics <- function(titre_est, timeSince, pars) {
    titre_est_new <- titre_est - (pars[1] * titre_est) * (timeSince)
    titre_est_new <- max(titre_est_new, 0)
    titre_est_new
}

infTuenisPower2016 <- function(titre_est, timeSince, pars) {

    y1 <- pars[1]
    t1 <- pars[2]
    r <- pars[3]
    alpha <- pars[4]

    v <- 0.001
    mu <- 1 / t1 * y1

    if (timeSince < t1) {
        titre_est_boost <- exp(mu * timeSince)
    } else {
        titre_est_boost <- exp(y1) * (1 + (r - 1) * exp(y1)^{r - 1} * v * (timeSince - t1)) ^ {-1 / (r - 1)}
    }

    titre_est_log <- titre_est + log(titre_est_boost) * max(0, 1 - titre_est * alpha)
    titre_est_log 
}


```

Create the antibody kinetics model 

```{r}

abkineticsModel <- list(
    model = makeModel(
        addAbkineticsModel("none", "IgG", "none", c("wane"), noInfSerumKinetics),
        addAbkineticsModel("inf", "IgG", "inf", c("y1_h1", "t1_h1", "r_h1", "s"), infTuenisPower2016)
    ),
    prior = bind_rows(
        add_par_df("y1_h1", 1, 6, "unif",  1, 6),
        add_par_df("t1_h1", 3, 14, "unif", 3, 14),
        add_par_df("r_h1", 1, 5, "unif", 1, 5),
        add_par_df("s", 0, 1, "unif", 0, 1),
        add_par_df("wane", 0, 1, "unif", 0, 1)
    )
)

```

- **`model = makeModel(...)`**: The antibody kinetics model is built using `makeModel`, which defines two different antibody response models.

- **`addAbkineticsModel("none", "PreF", "none", c("wane"), noInfSerumKinetics)`**: This line models the antibody kinetics when there is **no infection**. The "PreF" biomarker is modeled with the parameter `wane`, which represents the natural decay or waning of antibody levels over time. The `noInfSerumKinetics` function describes this process.

- **`addAbkineticsModel("inf", "PreF", "inf", c("y1_h1", "t1_h1", "r_h1", "s"), infTuenisPower2016)`**: This line models the antibody kinetics **after infection**. Several parameters describe the response to infection:
    - **`y1_h1`**: Initial antibody level post-infection.
    - **`t1_h1`**: Time of peak response.
    - **`r_h1`**: Rate of antibody decay post-infection.
    - **`s`**: Slope of decay after peak response.
    - **`infTuenisPower2016`**: The function governing the infection-induced antibody response.

- **`prior = bind_rows(...)`**: This block defines prior distributions for the model parameters:
    - **`y1_h1`**: Uniform prior between `1` and `6`, representing the initial antibody level after infection.
    - **`t1_h1`**: Normal prior with mean `14` and standard deviation `3`, representing the time to peak response.
    - **`r_h1`**: Uniform prior between `1` and `5`, representing the rate of antibody decay.
    - **`s`**: Uniform prior between `0` and `1`, representing the slope of the decay post-infection.
    - **`wane`**: Uniform prior between `0` and `1`, representing the natural waning of antibody levels over time.

### Summary

The antibody kinetics model accounts for the natural waning of antibodies in the absence of infection, as well as the antibody response after infection. Parameters such as `y1_h1`, `t1_h1`, and `r_h1` govern the kinetics of antibody production and decay. The model uses both uniform and normal priors to specify reasonable ranges for these parameters.


## 3. Full model definition

This R code defines a serological model using different components such as biomarkers, exposure types, observational and antibody kinetics models, and priors. The code creates a model definitions (`modeldefinition_e1_p1`) and uses these to create a serological model for data analysis (`seroModel_phirst_e1_p1`).

## Step-by-Step Breakdown

### Step 1: Define the First Model `modeldefinition`

```{r}

# See data_format.Rmd vignette about data input types 

biomarkers <- "IgG"

# Create the data_sero dataframe
data_sero <- data.frame(
  id = c(1, 1, 2, 2, 3, 3),
  time = c(1, 50, 1, 50, 1, 50),
  IgG = c(1.2, 4.4, 1.2, 4.4, 3.0, 3.0)
)

# Define possible exposure types
exposureTypes <- c("none", "inf")

# Create the exposure_data dataframe
exposure_data <- data.frame(
  id = c(1),
  time = c(14),
  exposure_type = c("inf")
)

exposureFitted <- "inf"

# Create the attack_rate_data dataframe
attack_rate_data <- data.frame(
  time = rep(1:25),
  prob = rep(1/25, 25)
)

modeldefinition <- list(
    biomarkers = biomarkers,
    exposureTypes = exposureTypes,
    exposureFitted = exposureFitted,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel
  #  exposurePrior = attack_rate_data,
   # exposurePriorType = "empirical"
)

seroModel <- createSeroJumpModel(data_sero, NULL, modeldefinition)

```


## 4. Run the model!

```{r}

save_info <- list(
    file_name = "simple",
    model_name = "ex1"
)

rj_settings <- list(
        numberChainRuns = 4, 
        iterations = 2000,
        burninPosterior = 1000,
        thin = 1
      #  runParallel = FALSE,
      #  onDebug = TRUE
)

model_summary <- runInfRJMCMC(seroModel, rj_settings, save_info = save_info)

```

## 5. Plot the outputs

```{r}

# Need to have save_info model_summary to run these
plotMCMCDiagnosis(model_summary, save_info = save_info) # plots the convergence diagnosis
plotPostFigs(model_summary, save_info = save_info) #plots some plots of the posteriors

```

`model_summary$post` has all the posterior distributions which the user can extract and plot themselves! 