---
title: "Case Study 1: Simulation Recovery"
author: "David Hodgson"
date: "`r Sys.Date()`"
output: html_document
---


# R Code Description

This document describes the steps and components of the provided R code for creating and running serological models using simulated data. The code defines observational models, antibody kinetics models, and runs inference using MCMC methods.

---

## Step 1: Load Required Libraries
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

library(devtools)
library(tidyr)
library(dplyr)
library(magrittr)
library(serojump)

```

## Step 2: Read Simulated Data
This data has been generated using `serosim`.

```{r}

modelname_sim <- "cesCOP_notd"
obs_er <- "0.5"

modeli <- readRDS(here::here("outputs", "sim_data", modelname_sim, "inputs.RDS"))
res <- readRDS(file = here::here("outputs", "sim_data", modelname_sim, paste0("sim_data_", obs_er, ".rds")))
data_titre_model <- res$observed_biomarker_states %>% select(i, t, value) %>% rename(id = i, time = t, titre = value) 

data_titre_model <- data_titre_model %>% mutate(biomarker = "sVNT") %>% as.data.frame %>% rename(sVNT = titre)


```

## Step 3: Define Likelihood and Kinetics Functions

```{r}

obsLogLikelihood <- function(titre_val, titre_est, pars) {
    ll <- dnorm(titre_val, titre_est, pars[1], log = TRUE)
}

infSerumKinetics <- function(titre_est, timeSince, pars) {
    a <- pars[1]
    b <- pars[2]
    c <- pars[3]
    if (timeSince < 14) {
        titre_est <- titre_est + log(exp(a) + exp(c)) * (timeSince) / 14;
    } else {
        titre_est <- titre_est + log(exp(a) * exp(-b/10 * (timeSince - 14)) + exp(c));
    }
    titre_est
}

noInfSerumKinetics <- function(titre_est, timeSince, pars) {
    titre_est_log <- titre_est - pars[1] * (timeSince)
    titre_est_log <- max(0, titre_est_log)
    titre_est_log
}

```

## Step 4: Define the Observational and Kinetics Models

```{r}


# Define the biomarkers and exposure types in the model
biomarkers <- c("sVNT")
exposureTypes <- c("none", "inf")
exposureFitted <- "inf"

# Define the observational model
observationalModel <- list(
    names = c("sVNT"),
    model = makeModel(
        addObservationalModel("sVNT", c("sigma"), obsLogLikelihood)
        ), # observational model,
    prior = bind_rows(
            add_par_df("sigma", 0.0001, 4, "unif", 0.0001, 4)
    )
)

# Define the antibody kinetics model
abkineticsModel <- list(
    model = makeModel(
            addAbkineticsModel("none", "sVNT", "none",  c("wane"), noInfSerumKinetics),
            addAbkineticsModel("inf", "sVNT", "inf", c("a", "b", "c"), infSerumKinetics)
        ),
    prior = bind_rows(
        add_par_df("wane", 0.0, 0.01, "unif", 0.0, 0.01), # observational model
        add_par_df("a", -6, 6, "norm",  2, 2), # ab kinetics
        add_par_df("b", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c", 0, 4, "unif", 0,  4) # ab kinetics 
    )
)

inf_prior <- function(N, E, I, K) {
    N_adj <- N - K
    E_adj <- E - K 
    logPriorExpInf <- lfactorial(E_adj) + lfactorial(N_adj - E_adj) - lfactorial(N_adj ) 
    logPriorExpInf
}

modeldefinition_cop <- list(
    biomarkers = biomarkers,
    exposureTypes = exposureTypes,
    exposureFitted = exposureFitted,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel,
    expInfPrior = inf_prior
)

model_cop <- createSeroJumpModel(data_titre_model, NULL, modeldefinition_cop)


```

## Step 5: Define the Complete Model


```{r}

modelname_sim <- "cesNoCOP_notd"
obs_er <- "0.5"

modeli <- readRDS(here::here("outputs", "sim_data", modelname_sim, "inputs.RDS"))
res <- readRDS(file = here::here("outputs", "sim_data", modelname_sim, paste0("sim_data_", obs_er, ".rds")))
data_titre_model <- res$observed_biomarker_states %>% select(i, t, value) %>% rename(id = i, time = t, titre = value) 

data_titre_model <- data_titre_model %>% mutate(biomarker = "sVNT") %>% as.data.frame %>% rename(sVNT = titre)


```

```{r}

obsLogLikelihood <- function(titre_val, titre_est, pars) {
    ll <- dnorm(titre_val, titre_est, pars[1], log = TRUE)
}

infSerumKinetics <- function(titre_est, timeSince, pars) {
    a <- pars[1]
    b <- pars[2]
    c <- pars[3]
    if (timeSince < 14) {
        titre_est <- titre_est + log(exp(a) + exp(c)) * (timeSince) / 14;
    } else {
        titre_est <- titre_est + log(exp(a) * exp(-b/10 * (timeSince - 14)) + exp(c));
    }
    titre_est
}

noInfSerumKinetics <- function(titre_est, timeSince, pars) {
    titre_est_log <- titre_est - pars[1] * (timeSince)
    titre_est_log <- max(0, titre_est_log)
    titre_est_log
}

```


```{r}


# Define the biomarkers and exposure types in the model
biomarkers <- c("sVNT")
exposureTypes <- c("none", "inf")
exposureFitted <- "inf"

# Define the observational model
observationalModel <- list(
    names = c("sVNT"),
    model = makeModel(
        addObservationalModel("sVNT", c("sigma"), obsLogLikelihood)
        ), # observational model,
    prior = bind_rows(
            add_par_df("sigma", 0.0001, 2, "exp", 1, NA)
    )
)

# Define the antibody kinetics model
abkineticsModel <- list(
    model = makeModel(
            addAbkineticsModel("none", "sVNT", "none",  c("wane"), noInfSerumKinetics),
            addAbkineticsModel("inf", "sVNT", "inf", c("a", "b", "c"), infSerumKinetics)
        ),
    prior = bind_rows(
        add_par_df("wane", 0.0, 0.01, "unif", 0.0, 0.01), # observational model
        add_par_df("a", -6, 6, "norm",  2, 2), # ab kinetics
        add_par_df("b", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c", 0, 4, "unif", 0,  4) # ab kinetics 
    )
)

inf_prior <- function(N, E, I, K) {
    N_adj <- N - K
    E_adj <- E - K 
    logPriorExpInf <- lfactorial(E_adj) + lfactorial(N_adj - E_adj) - lfactorial(N_adj ) 
    logPriorExpInf
}



modeldefinition_no_cop <- list(
    biomarkers = biomarkers,
    exposureTypes = exposureTypes,
    exposureFitted = exposureFitted,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel,
    expInfPrior = inf_prior
)

model_no_cop <- createSeroJumpModel(data_titre_model, NULL, modeldefinition_no_cop)


```


```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

rj_settings <- list(
        numberChainRuns = 4, 
        iterations = 400000,
        burninPosterior = 200000,
        thin = 1000
    )

save_info_cop <- list(
    file_name = "simulated_data",
    model_name = "cop"
)

save_info_no_cop <- list(
    file_name = "simulated_data",
    model_name = "no_cop"
)

# Run these but take a while
#model_summary_cop <- runInfRJMCMC(model_cop, rj_settings, save_info = save_info_cop)
#model_summary_no_cop <- runInfRJMCMC(model_no_cop, rj_settings, save_info = save_info_no_cop)

model_summary_cop <-  readRDS(here::here("outputs", "fits", save_info_cop$file_name, save_info_cop$model_name, paste0("model_summary.RDS")))
model_summary_no_cop <-  readRDS(here::here("outputs", "fits", save_info_no_cop$file_name, save_info_no_cop$model_name, paste0("model_summary.RDS")))

modelname_sim <- "cesCOP_notd"
sim_model_cop <- readRDS(here::here("outputs", "sim_data", modelname_sim, "inputs.RDS"))
sim_res_cop <- readRDS(file = here::here("outputs", "sim_data", modelname_sim, paste0("sim_data_", "0.5", ".rds")))

plotMCMCDiagnosis(model_summary_cop, save_info = save_info_cop)
plotPostFigsSim(model_summary_cop, sim_model_cop, sim_res_cop, save_info = save_info_cop)

modelname_sim <- "cesNoCOP_notd"
sim_model_no_cop <- readRDS(here::here("outputs", "sim_data", modelname_sim, "inputs.RDS"))
sim_res_no_cop <- readRDS(file = here::here("outputs", "sim_data", modelname_sim, paste0("sim_data_", "0.5", ".rds")))

plotMCMCDiagnosis(model_summary_no_cop, save_info = save_info_no_cop)
plotPostFigsSim(model_summary_no_cop, sim_model_no_cop, sim_res_no_cop, save_info = save_info_no_cop)

```