
Fitting to the TRANSVIR data prototype

# Load the required packages

```{r}

#install.packages("devtools")
#library(devtools)
devtools::load_all()

library(serojump)
library(lubridate)
library(patchwork)
library(tidybayes)
library(ggdist)

```


# Describe functions required for the rjmcmc model

```{r}


obsLogLikelihoodSerum <- function(titre_val, titre_est, pars) {

    if (titre_val <= log10(40)) {
        ll <- dnorm(titre_val, titre_est, pars[1], log = TRUE)#pnorm(log10(40), titre_est, pars[1], log.p = TRUE)
    } else {
        ll <- dnorm(titre_val, titre_est, pars[1], log = TRUE)
    }
    ll
}

obsLogLikelihoodIgA <- function(titre_val, titre_est, pars) {
    ll <- dnorm(titre_val, titre_est, pars[1], log = TRUE)
}


noInfSerumKinetics <- function(titre_est, timeSince, pars) {
    titre_est_log <- titre_est - pars[1] * (timeSince)
    titre_est_log <- max(0, titre_est_log)
    titre_est_log
}


infTuenisPower2016 <- function(titre_est, timeSince, pars) {

    y1 <- pars[1]
    t1 <- pars[2]
    r <- pars[3]
    alpha <- pars[4]

    v <- 0.001
    mu <- 1 / t1 * y1

    if (timeSince < t1) {
        titre_est_boost <- exp(mu * timeSince)
    } else {
        titre_est_boost <- exp(y1) * (1 + (r - 1) * exp(y1)^{r - 1} * v * (timeSince - t1)) ^ {-1 / (r - 1)}
    }

    titre_est_log <- titre_est + log(titre_est_boost) * max(0, 1 - titre_est * alpha)
    titre_est_log
}


```

# Delta wave
## Load cleaned data

```{r}

gambia_pvnt_w2 <- get_data_titre_model_wave2_pvnt_iga()# This is the empirical prior for the exposure time
gambia_exp_w2 <- get_exposures_wave2() # This is the empirical prior for the exposure time
exp_prior_w2 <- get_exp_prior_wave2() # This is the empirical prior for the exposure time


known_exposure_gambia_exp_w2_add <- gambia_exp_w2 %>% mutate(type = "known")
sero_confirmed <-  gambia_pvnt_w2 %>% group_by(id) %>% mutate(r = row_number(), r_max = max(r)) %>% filter(r_max >=2) %>%
        select(!c(IgA, sVNT, time )) %>% pivot_wider(names_from = r, values_from = spike) %>% 
        mutate(diff = `1` - `2`) %>% filter(diff < -0.602) %>% left_join(known_exposure_gambia_exp_w2_add) %>% 
        filter(is.na(type)) 
rate_inf_w2 <- (sero_confirmed %>% nrow) / (gambia_pvnt_w2$id %>% unique %>% length)

```



## Describe the observational and antibody kinetics model

```{r}



# Define the biomarkers and exposure types in the model
biomarkers <- c("spike", "IgA")
exposureTypes <- c("none", "delta", "vax", "predelta")
exposureFitted <- "delta"

# Define the observational model
observationalModel <- list(
    names = c("spike", "IgA"),
    model = makeModel(
        addObservationalModel("spike", c("sigma"), obsLogLikelihoodSerum),
        addObservationalModel("IgA", c("sigma_a"), obsLogLikelihoodIgA)),
    prior = bind_rows(
        add_par_df("sigma", 0.0001, 2, "exp", 1, NA),
        add_par_df("sigma_a", 0.0001, 2, "exp", 1, NA)
        ) # observational model,
)

# Define the antibody kinetics model
abkineticsModel <- list(
    model = makeModel(
            addAbkineticsModel("none", "spike", "none",  c("wane"), noInfSerumKinetics),
            addAbkineticsModel("delta", "spike", "delta", c("y1_d", "t1_d", "r_d", "s"), infTuenisPower2016),
            addAbkineticsModel("vax", "spike", "vax", c("y1_vax", "t1_vax", "r_vax", "s"), infTuenisPower2016),
            addAbkineticsModel("predelta", "spike", "predelta", c("y1_pd", "t1_pd", "r_pd", "s"), infTuenisPower2016),
            addAbkineticsModel("none_a", "IgA", "none",  c("wane_a"), noInfSerumKinetics),
            addAbkineticsModel("delta_a", "IgA", "delta", c("y1_d_a", "t1_d_a", "r_d_a", "s_a"), infTuenisPower2016),
            addAbkineticsModel("vax_a", "IgA", "vax", c("y1_vax_a", "t1_vax_a", "r_vax_a", "s_a"), infTuenisPower2016),
            addAbkineticsModel("predelta_a", "IgA", "predelta", c("y1_pd_a", "t1_pd_a", "r_pd_a", "s_a"), infTuenisPower2016)
        ),
    prior = bind_rows(
        add_par_df("y1_d", 1, 6, "unif",  1, 6), # ab kinetics
        add_par_df("t1_d", 10, 25, "unif",  10, 25), # ab kinetics
        add_par_df("r_d", 1, 5, "unif", 1, 5), # ab kinetics 
        add_par_df("y1_vax", 1, 6, "unif",  1, 6), # ab kinetics
        add_par_df("t1_vax", 10, 25, "unif",  10, 25), # ab kinetics
        add_par_df("r_vax", 1, 5, "unif", 1, 5), # ab kinetics 
        add_par_df("y1_pd", 1, 9, "unif",  1, 9), # ab kinetics
        add_par_df("t1_pd", 10, 25, "unif",  10, 25), # ab kinetics
        add_par_df("r_pd", 1, 7, "unif", 1, 7), # ab kinetics 
        add_par_df("s",  0, 1, "unif", 0, 1), # ab kinetics 
        add_par_df("wane", 0.0, 0.01, "unif", 0.0, 0.01), # observational model
        add_par_df("y1_d_a", 1, 6, "unif",  1, 6), # ab kinetics
        add_par_df("t1_d_a", 25, 75, "unif",  25, 75), # ab kinetics
        add_par_df("r_d_a", 1, 5, "unif", 1, 5), # ab kinetics 
        add_par_df("y1_vax_a", 1, 6, "unif",  1, 6), # ab kinetics
        add_par_df("t1_vax_a", 25, 75, "unif",  25, 75), # ab kinetics
        add_par_df("r_vax_a", 1, 5, "unif", 1, 5), # ab kinetics 
        add_par_df("y1_pd_a", 1, 6, "unif",  1, 6), # ab kinetics
        add_par_df("t1_pd_a", 25, 75, "unif",  25, 75), # ab kinetics
        add_par_df("r_pd_a", 1, 7, "unif", 1, 7), # ab kinetics 
        add_par_df("s_a", 0, 1, "unif", 0, 1), # ab kinetics 
        add_par_df("wane_a", 0.0, 0.01, "unif", 0.0, 0.01) # observational model
    )
)


```

## Define priors on the number of infected person 

```{r}


inf_prior_1 <- function(N, E, I, K) {
    0
}

inf_prior_2 <- function(N, E, I, K) {
    N_adj <- N - K
    E_adj <- E - K 
    logPriorExpInf <- lfactorial(E_adj) + lfactorial(N_adj - E_adj) - lfactorial(N_adj ) 
    logPriorExpInf
}

inf_prior_3 <- function(N, E, I, K) {
    N_adj <- N - K
    E_adj <- E - K
    logPriorExpInf <- lfactorial(E_adj) + lfactorial(N_adj - E_adj) - lfactorial(N_adj ) + dbinom(E_adj, N_adj, 0.117, log = TRUE)
   # cat(logPriorExpInf, "\n")
    logPriorExpInf
}


```

## Create model for serojump

```{r}

modeldefinition_w2_p1 <- list(
    biomarkers = biomarkers,
    exposureTypes = exposureTypes,
    exposureFitted = exposureFitted,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel,
    exposurePrior = exp_prior_w2,
    exposurePriorType = "empirical",
    expInfPrior = inf_prior_1
)


modeldefinition_w2_p2 <- modeldefinition_w2_p1
modeldefinition_w2_p2$expInfPrior <- inf_prior_2

modeldefinition_w2_p3 <- modeldefinition_w2_p1
modeldefinition_w2_p3$expInfPrior <- inf_prior_3


modelW2_p1 <- createSeroJumpModel(gambia_pvnt_w2, gambia_exp_w2, modeldefinition_w2_p1, TRUE)
modelW2_p2 <- createSeroJumpModel(gambia_pvnt_w2, gambia_exp_w2, modeldefinition_w2_p2)
modelW2_p3 <- createSeroJumpModel(gambia_pvnt_w2, gambia_exp_w2, modeldefinition_w2_p3)


```

# Omicron wave
## Load cleaned data

```{r}


gambia_pvnt_w3 <- get_data_titre_model_wave3_pvnt_iga()# This is the empirical prior for the exposure time
gambia_exp_w3 <- get_exposures_wave3() # This is the empirical prior for the exposure time
exp_prior_w3 <- get_exp_prior_wave3() # This is the empirical prior for the exposure time


known_exposure_gambia_exp_w3_add <- gambia_exp_w3 %>% mutate(type = "known")
sero_confirmed <-  gambia_pvnt_w3 %>% group_by(id) %>% mutate(r = row_number(), r_max = max(r)) %>% filter(r_max >=2) %>%
        select(!c(IgA, time, sVNT )) %>% pivot_wider(names_from = r, values_from = spike) %>% 
        mutate(diff = `1` - `2`) %>% filter(diff < -0.602) %>% left_join(known_exposure_gambia_exp_w3_add) %>% 
        filter(is.na(type))
# -0.602 4-fold rise
# -0.903 8-fold rise

rate_inf_w3 <- (sero_confirmed %>% nrow) / (gambia_pvnt_w3$id %>% unique %>% length)

```



## Describe the observational and antibody kinetics model

```{r}



# Define the biomarkers and exposure types in the model
biomarkers <- c("spike", "IgA")
exposureTypes <- c("none", "omicron", "vax")
exposureFitted <- "omicron"

# Define the observational model
observationalModel <- list(
    names = c("spike", "IgA"),
    model = makeModel(
        addObservationalModel("spike", c("sigma"), obsLogLikelihoodSerum),
        addObservationalModel("IgA", c("sigma_a"), obsLogLikelihoodIgA)),

    prior = bind_rows(
        add_par_df("sigma", 0.0001, 2, "exp", 1, NA),
        add_par_df("sigma_a", 0.0001, 2, "exp", 1, NA)
        ) # observational model,
)

# Define the antibody kinetics model
abkineticsModel <- list(
    model = makeModel(
            addAbkineticsModel("none", "spike", "none",  c("wane"), noInfSerumKinetics),
            addAbkineticsModel("omicron", "spike", "omicron", c("y1_o", "t1_o", "r_o", "s"), infTuenisPower2016),
            addAbkineticsModel("vax", "spike", "vax", c("y1_vax", "t1_vax", "r_vax", "s"), infTuenisPower2016),
            addAbkineticsModel("none_a", "IgA", "none",  c("wane_a"), noInfSerumKinetics),
            addAbkineticsModel("omicron_a", "IgA", "omicron", c("y1_o_a", "t1_o_a", "r_o_a", "s_a"), infTuenisPower2016),
            addAbkineticsModel("vax_a", "IgA", "vax", c("y1_vax_a", "t1_vax_a", "r_vax_a", "s_a"), infTuenisPower2016)
        ),
    prior = bind_rows(
        add_par_df("y1_o", 1, 6, "unif",  1, 6), # ab kinetics
        add_par_df("t1_o", 10, 25, "unif",  10, 25), # ab kinetics
        add_par_df("r_o", 1, 5, "unif", 1, 5), # ab kinetics 
        add_par_df("y1_vax", 1, 9, "unif",  1, 9), # ab kinetics
        add_par_df("t1_vax", 10, 25, "unif",  10, 25), # ab kinetics
        add_par_df("r_vax", 1, 5, "unif", 1, 5), # ab kinetics 
        add_par_df("s",  0, 1, "unif", 0, 1), # ab kinetics 
        add_par_df("wane", 0.0, 0.01, "unif", 0.0, 0.01), # observational model
        add_par_df("y1_o_a", 1, 6, "unif",  1, 6), # ab kinetics
        add_par_df("t1_o_a", 25, 75, "unif",  25, 75), # ab kinetics
        add_par_df("r_o_a", 1, 5, "unif", 1, 5), # ab kinetics 
        add_par_df("y1_vax_a", 1, 6, "unif",  1, 6), # ab kinetics
        add_par_df("t1_vax_a", 25, 75, "unif",  25, 75), # ab kinetics
        add_par_df("r_vax_a", 1, 5, "unif", 1, 5), # ab kinetics 
        add_par_df("s_a", 0, 1, "unif", 0, 1), # ab kinetics 
        add_par_df("wane_a", 0.0, 0.01, "unif", 0.0, 0.01) # observational model
    )
)

```

## Define priors on the number of infected person 

```{r}


inf_prior_1 <- function(N, E, I, K) {
    0
}

inf_prior_2 <- function(N, E, I, K) {
    N_adj <- N - K
    E_adj <- E - K 
    logPriorExpInf <- lfactorial(E_adj) + lfactorial(N_adj - E_adj) - lfactorial(N_adj ) 
    logPriorExpInf
}

inf_prior_3 <- function(N, E, I, K) {
    N_adj <- N - K
    E_adj <- E - K
    logPriorExpInf <- lfactorial(E_adj) + lfactorial(N_adj - E_adj) - lfactorial(N_adj ) + dbinom(E_adj, N_adj,  0.122807, log = TRUE)
    N_adj <- N - K
  #  cat(logPriorExpInf, "\n")
    logPriorExpInf
}

```

## Create model for serojump

```{r}

modeldefinition_w3_p1 <- list(
    biomarkers = biomarkers,
    exposureTypes = exposureTypes,
    exposureFitted = exposureFitted,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel,
    exposurePrior = exp_prior_w3,
    exposurePriorType = "empirical",
    expInfPrior = inf_prior_1
)

modeldefinition_w3_p2 <- modeldefinition_w3_p1
modeldefinition_w3_p2$expInfPrior <- inf_prior_2

modeldefinition_w3_p3 <- modeldefinition_w3_p1
modeldefinition_w3_p3$expInfPrior <- inf_prior_3


modelW3_p1 <- createSeroJumpModel(gambia_pvnt_w3, gambia_exp_w3, modeldefinition_w3_p1, TRUE)
modelW3_p2 <- createSeroJumpModel(gambia_pvnt_w3, gambia_exp_w3, modeldefinition_w3_p2)
modelW3_p3 <- createSeroJumpModel(gambia_pvnt_w3, gambia_exp_w3, modeldefinition_w3_p3)


```

# Fit the model using the rjmcmc packages

```{r}

rj_settings <- list(
        numberChainRuns = 4, 
        iterations = 2000,
        burninPosterior = 1000,
        thin = 10
    )

save_info <- list(
    file_name = "transvir_data",
    model_name = "wave3_p3"
)

model_summary_p3_w3 <- runInfRJMCMC(modelW3_p3, rj_settings, save_info = save_info, priorPred = TRUE)


# Need to have save_info defined to run these
plotMCMCDiagnosis(model_summary_p3_w3, save_info = save_info)
plotPostFigs(model_summary_p3_w3, save_info = save_info)

```






rj_settings <- list(
        numberChainRuns = 4, 
       # numberCores = 4,
        iterations = 2000,
        burninPosterior = 1000,
        thin = 10
       # consoleUpdates = 100,
       # onAdaptiveCov = TRUE,
       # updatesAdaptiveCov = 10,
        #burninAdaptiveCov = 1000
        #covarInitVal = 1e-2, # make very small if struggling to sample to beginning
        #covarInitValAdapt = 1e-2, # make very small if struggling to sample to beginning
        #covarMaxVal = 1, # decrease if struggling toc sample in the middle
        #runParallel = FALSE
     #   noGibbsSteps = 1
        #noGibbsSteps = 1,
        #onDebug = FALSE
    )