Fitting to the CRICK data

# Load the data
```{r}

#install.packages("devtools")
#library(devtools)
devtools::load_all()

library(lubridate)
library(patchwork)
library(tidybayes)
library(ggdist)

```

## Get data in the right format for rjmcmc model

To help with this section please consult information in the documentation on how to layour this data.frame:

* For serological data in observational model see [data_titre](../man/data_titre.Rd).
* For known exposure times [data_known_exposures](./man/data_known_exposures.Rd).
* For exposure prior see XXX.
  
```{r}
# Load data clean in the formet required for the model
# THESE FUNCTIONS NEED TO CHANGE TO LOAD IN THE CRICK DATA

data_exp <- get_data_titre_nih_2023()
data_titre <- data_exp[[1]] #%>% check_titre
known_exposure <- data_exp[[2]] %>% unique # must be unique
exp_prior <- data_exp[[3]]

data_titre %>% check_titre

```

```{r}
# Define the functions 

obsLogLikelihood <- function(titre_val, titre_est, pars) {
    if (titre_val < 0) {
        ll <- pnorm(0, titre_est, pars[1], log.p = TRUE)
    } else if (titre_val >= 8){
        ll <- 1 - pnorm(8, titre_est, pars[1], log.p = TRUE)
    } else {
        ll <- log(pnorm(titre_val + 1, titre_est, pars[1]) - pnorm(titre_val , titre_est, pars[1]))
    }
    ll
}

copLogLikelihood <- function(inf_status, esttitreExp, params) {
    # COP parameters
    beta0 <- params[1]
    beta1 <- params[2]
    p <- 1.0 / (1.0 + exp(- (beta0 + beta1 * esttitreExp) ) )
    ll <- inf_status * log(p) + (1 - inf_status) * log(1 - p)
    ll
}

noInfSerumKinetics <- function(titre_est, timeSince, pars) {
    titre_est <- titre_est - pars[1] * (timeSince)
    titre_est <- max(titre_est, 0)
    titre_est
}

infSerumKinetics <- function(titre_est, timeSince, pars) {
    a <- pars[1]
    b <- pars[2]
    c <- pars[3]
    if (timeSince < 14) {
        titre_est <- titre_est + log(exp(a) + exp(c)) * (timeSince) / 14;
    } else {
        titre_est <- titre_est + log(exp(a) * exp(-b/10 * (timeSince - 14)) + exp(c));
    }
    titre_est
}

```



## Define the models for the rjmcmc model

To help with this section please consult information in the documentation on how to define these lists:

* For observation model see [observationalModel](./man/observationalModel.Rd).
* For COP model see [copModel](./man/copModel.Rd).
* For antibody kinetics see [abkineticsModel](./man/abkineticsModel.Rd).

```{r}


# Define the biomarkers and exposure types in the model
biomakers <- c("A/Darwin/06/2021")
exposureType <- c("vax", "h3_2023")

# Define the observational model
observationalModel <- list(
    names = c("A/Darwin/06/2021"),
    model = makeModel(addObservationalModel("A/Darwin/06/2021", c("sigma"), obsLogLikelihood)),
    prior = add_par_df("sigma", 0.0001, 4, "unif", 0.0001, 4) # observational model,
)

# Define the antibody kinetics model
abkineticsModel <- list(
    names = c("vax", "h3_2023"),
    model = makeModel(
            addAbkineticsModel("A/Darwin/06/2021", "vax", FALSE, c("a_vax", "b_vax", "c_vax"), infSerumKinetics),
            addAbkineticsModel("A/Darwin/06/2021", "h3_2023", TRUE, c("a_v", "b_v", "c_v"), infSerumKinetics)
        ),
    prior = bind_rows(
        add_par_df("a_vax", -2, 2, "norm",  0, 1), # ab kinetics
        add_par_df("b_vax", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c_vax", 0, 4, "unif", 0,  4), # ab kinetics 
        add_par_df("a_v", -2, 2, "norm",  0, 1), # ab kinetics
        add_par_df("b_v", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c_v", 0, 4, "unif", 0,  4), # ab kinetics
        add_par_df("wane", 0.0, 0.1, "unif", 0.0, 0.1) # observational model
    )
)

# Define the COP model
copModel <- list( 
        names = c("A/Darwin/06/2021"),
        model = makeModel(addCopModel("A/Darwin/06/2021", "h3_2023", c("beta0", "beta1"), copLogLikelihood)),
        prior = bind_rows(
            add_par_df("beta0", -10, 10, "unif", -10, 10), # cop model (not used here)
            add_par_df("beta1", -10, 10, "unif", -10, 10)
        ) # cop model (not used here),
)

modeldefinition <- list(
    biomarkers = biomakers,
    exposureType = exposureType,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel,
    copModel = copModel,
    exposurePrior = exp_prior,
    exposurePriorType = "empirical"
)

modelW2 <- createSeroJumpModel(data_titre, known_exposure, modeldefinition)

````


```{r}

# Define the settings
settings <-  list(
    numberChainRuns = 4,
    numberCores = 4,
    iterations = 200000,
    burninPosterior = 100000,
    thin = 100,
    consoleUpdates = 100,
    onAdaptiveCov = TRUE,
    updatesAdaptiveCov = 100,
    burninAdaptiveCov = 1000,
    covarInitVal = 1e-2, # make very small if struggling to sample to beginning
    covarInitValAdapt = 1e-2, # make very small if struggling to sample to beginning
    covarMaxVal = 1, # decrease if struggling to sample in the middle
    runParallel = TRUE,
    noGibbsSteps = 1,
    onDebug = FALSE
)

# Run the model 
runRJMCMC(modelW2, settings, "test/nih_2024", "h3")

```

# Profiling 

```{r}

devtools::load_all()

settings <-  list(
    numberChainRuns = 1,
    numberCores = 4,
    iterations = 100,
    burninPosterior = 50,
    thin = 1,
    consoleUpdates = 10,
    onAdaptiveCov = TRUE,
    updatesAdaptiveCov = 10,
    burninAdaptiveCov = 1000,
    covarInitVal = 1e-2, # make very small if struggling to sample to beginning
    covarInitValAdapt = 1e-2, # make very small if struggling to sample to beginning
    covarMaxVal = 1, # decrease if struggling to sample in the middle
    runParallel = FALSE,
    noGibbsSteps = 1,
    onDebug = FALSE,
    profile = TRUE
)

runRJMCMC(modelW2, settings, "test/nih_2024", "h3_profile")

library(RcppClock)
profiletimes

```