Get started: a very simple working example 

## Load the packages

```{r}

#install.packages("devtools")
#library(devtools)
devtools::load_all()
library(rjmc)
library(lubridate)
library(patchwork)
library(tidybayes)
library(ggdist)

```

## Get data in the right format for rjmcmc model


To help with this section please consult information in the documentation on how to layour this data.frame:

* For serological data in observational model see [data_titre](../man/data_titre.Rd).
* For known exposure times [data_known_exposures](./man/data_known_exposures.Rd).
* For exposure prior see XXX.

```{r}

# Load data clean in the format required for the model
data_titre <- data.frame(
    id = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
    time = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
    titre = c(1.5, 1.6, 1.7, 1.8, 1.9, 2.0, 2.1, 2.2, 2.3, 2.4),
    biomarker = rep("IgG", 10)
 )

# No known exposure times in this model 
data_known_exposures <- NULL

exp_prior_w2 <- get_exp_prior_wave2() # This is the empirical prior for the exposure time

```

## Define the functions required for the rjmcmc model

To help with this section please consult information in the documentation on how to define these functions:

* For observation functions see [obsLogLikelihood](./man/obsLogLikelihood.Rd).
* For COP model functions [copLogLikelihood](./man/copLogLikelihood.Rd).
* For antibody kinetics functions see [abkineticsFunction](./man/abkineticsFunction.Rd).

```{r}

# This function uses a cauchy distribution for the likelihood of the titre values with a LOD a titre value ofg log10(40)
obsLogLikelihood = function(titre_val, titre_est, pars) {
    if (titre_val <= log10(40)) {
        ll <- pcauchy(log10(40), titre_est, pars[1], log.p = TRUE)
    } else {
        ll <- dcauchy(titre_val, titre_est, pars[1], log = TRUE)
    }
    ll
}

# This function uses a logistic distribution for the likelihood.
copLogLikelihood <- function(inf_status, esttitreExp, params) {
    # COP parameters
    beta0 <- params[1]
    beta1 <- params[2]
    p <- 1.0 / (1.0 + exp(- (beta0 + beta1 * esttitreExp) ) )
    ll <- inf_status * log(p) + (1 - inf_status) * log(1 - p)
    ll
}

# This function uses describes the antibody kinetics model for the time until infection
noInfSerumKinetics <- function(titre_est, timeSince, pars) {
    titre_est <- titre_est - pars[1] * (timeSince)
    titre_est <- max(0, titre_est)
    titre_est
}

# This function uses describes the antibody kinetics model after infection
infSerumKinetics <- function(titre_est, timeSince, pars) {
    a <- pars[1] 
    b <- pars[2] 
    c <- pars[3]
    if (timeSince < 14) {
        titre_est <- titre_est + log(exp(a) + exp(c)) * (timeSince) / 14;
    } else {
        titre_est <- titre_est + log(exp(a) * exp(-b/10 * (timeSince - 14)) + exp(c));
    }
    titre_est
}

```


## Define the models for the rjmcmc model

To help with this section please consult information in the documentation on how to define these lists:

* For observation model see [observationalModel](./man/observationalModel.Rd).
* For COP model see [copModel](./man/copModel.Rd).
* For antibody kinetics see [abkineticsModel](./man/abkineticsModel.Rd).

```{r}


# Define the biomarkers and exposure types in the model
biomakers <- c("IgG")
exposureType <- c("delta", "vax", "predelta")

# Define the observational model
observationalModel <- list(
    names = c("IgG"),
    model = makeModel(addObservationalModel("IgG", c("sigma"), obsLogLikelihood)),
    prior = add_par_df("sigma", 0.0001, 4, "unif", 0.0001, 4) # observational model,
)

# Define the antibody kinetics model
abkineticsModel <- list(
    names = c("none", "delta", "predelta", "vax"),
    model = makeModel(
            addAbkineticsModel("IgG", "none", FALSE,  c("wane"), noInfSerumKinetics),
            addAbkineticsModel("IgG", "delta", TRUE, c("a_d", "b_d", "c_d"), infSerumKinetics),
            addAbkineticsModel("IgG", "predelta", FALSE, c("a_pd", "b_pd", "c_pd"), infSerumKinetics),
            addAbkineticsModel("IgG", "vax", FALSE, c("a_vax", "b_vax", "c_vax"), infSerumKinetics)
        ),
    prior = bind_rows(
        add_par_df("a_vax", -2, 2, "norm",  0, 1), # ab kinetics
        add_par_df("b_vax", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c_vax", 0, 4, "unif", 0,  4), # ab kinetics 
        add_par_df("a_pd", -2, 2, "norm",  0, 1), # ab kinetics
        add_par_df("b_pd", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c_pd", 0, 4, "unif", 0,  4), # ab kinetics
        add_par_df("a_d", -2, 2, "norm",  0, 1), # ab kinetics
        add_par_df("b_d", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c_d", 0, 4, "unif", 0,  4), # ab kinetics
        add_par_df("wane", 0.0, 0.1, "unif", 0.0, 0.1) # observational model
    )
)

# Define the COP model
copModel <- list( 
        names = c("IgG"),
        model = makeModel(addCopModel("IgG", "delta", c("beta0", "beta1"), copLogLikelihood)),
        prior = bind_rows(
            add_par_df("beta0", -10, 10, "unif", -10, 10), # cop model (not used here)
            add_par_df("beta1", -10, 10, "unif", -10, 10)
        ) # cop model (not used here),
)

modeldefinition <- list(
    biomarkers = biomakers,
    exposureType = exposureType,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel,
    copModel = copModel,
    exposurePrior = exp_prior_w2,
    exposurePriorType = "empirical"
)

modelW2 <- createSeroJumpModel(data_titre, data_known_exposures, modeldefinition)

````

# Define settings and run the model, (this won't run as made the data up)

```{r eval = FALSE}

# Define the settings
settings <-  list(
    numberChainRuns = 4,
    numberCores = 4,
    iterations = 200000,
    burninPosterior = 100000,
    thin = 100,
    consoleUpdates = 100,
    onAdaptiveCov = TRUE,
    updatesAdaptiveCov = 100,
    burninAdaptiveCov = 1000,
    updatesAdaptiveTemp = 10,
    covarInitVal = 1e-2, # make very small if struggling to sample to beginning
    covarInitValAdapt = 1e-2, # make very small if struggling to sample to beginning
    covarMaxVal = 1, # decrease if struggling to sample in the middle
    runParallel = FALSE,
    noGibbsSteps = 1
)

# Run the model 
runRJMCMC(modelW2, settings, "test/transvir", "test")

```