

Fitting to the TRANSVIR data prototype

# Load the data

```{r}

#install.packages("devtools")
#library(devtools)
devtools::load_all()

library(lubridate)
library(patchwork)
library(tidybayes)
library(ggdist)

```

## Get data in the right format for rjmcmc model

To help with this section please consult information in the documentation on how to layour this data.frame:

* For serological data in observational model see [data_titre](../man/data_titre.Rd).
* For known exposure times [data_known_exposures](./man/data_known_exposures.Rd).
* For exposure prior see XXX.
  
```{r}

# 41-483J
# 43-512H
gambia_pvnt_w1 <- get_data_titre_model_wave1() %>% rename(IgG = titre)# This is the empirical prior for the exposure time


# Load data clean in the formet required for the model
gambia_pvnt_w2 <- get_data_titre_model_wave2() %>% rename(IgG = titre)# This is the empirical prior for the exposure time
gambia_exp_w2 <- get_exposures_wave2() # This is the empirical prior for the exposure time
exp_prior_w2 <- get_exp_prior_wave2() # This is the empirical prior for the exposure time

gambia_pvnt_w2 %>% left_join(gambia_exp_w2 %>% rename(exp_time = time))
```

## Define the functions required for the rjmcmc model

To help with this section please consult information in the documentation on how to define these functions:

* For observation functions see [obsLogLikelihood](./man/obsLogLikelihood.Rd).
* For COP model functions [copLogLikelihood](./man/copLogLikelihood.Rd).
* For antibody kinetics functions see [abkineticsFunction](./man/abkineticsFunction.Rd).


```{r}
# Define the functions 

obsLogLikelihood = function(titre_val, titre_est, pars) {
    if (titre_val <= log10(40)) {
        ll <- pcauchy(log10(40), titre_est, pars[1], log.p = TRUE)
    } else {
        ll <- dcauchy(titre_val, titre_est, pars[1], log = TRUE)
    }
    ll
}

copFuncForm <- function(inf_status, esttitreExp, params) {
    beta0 <- params[1]
    beta1 <- params[2]
    p <- 1.0 / (1.0 + exp(- (beta0 + beta1 * esttitreExp) ) )
}

copLogLikelihood <- function(inf_status, esttitreExp, params) {
    # COP parameters
    beta0 <- params[1]
    beta1 <- params[2]
    p <- 1.0 / (1.0 + exp(- (beta0 + beta1 * esttitreExp) ) )
    ll <- inf_status * log(p) + (1 - inf_status) * log(1 - p)
    ll
}

noInfSerumKinetics <- function(titre_est, timeSince, pars) {
    titre_est <- titre_est - pars[1] * (timeSince)
    titre_est
}

infSerumKinetics <- function(titre_est, timeSince, pars) {
    a <- pars[1]
    b <- pars[2]
    c <- pars[3]
    if (timeSince < 14) {
        titre_est <- titre_est + log(exp(a) + exp(c)) * (timeSince) / 14;
    } else {
        titre_est <- titre_est + log(exp(a) * exp(-b/10 * (timeSince - 14)) + exp(c));
    }
    titre_est
}

```

## Define the models for the rjmcmc model

To help with this section please consult information in the documentation on how to define these lists:

* For observation model see [observationalModel](./man/observationalModel.Rd).
* For COP model see [copModel](./man/copModel.Rd).
* For antibody kinetics see [abkineticsModel](./man/abkineticsModel.Rd).

```{r}


# Define the biomarkers and exposure types in the model
biomarkers <- c("IgG")
exposureTypes <- c("none", "delta", "vax", "predelta")
exposureFitted <- "delta"

# Define the observational model
observationalModel <- list(
    names = c("IgG"),
    model = makeModel(addObservationalModel("IgG", c("sigma"), obsLogLikelihood)),
    prior = add_par_df("sigma", 0.0001, 4, "unif", 0.0001, 4) # observational model,
)

# Define the antibody kinetics model
abkineticsModel <- list(
    model = makeModel(
            addAbkineticsModel("none", "IgG", "none",  c("wane"), noInfSerumKinetics),
            addAbkineticsModel("delta", "IgG", "delta", c("a_d", "b_d", "c_d"), infSerumKinetics),
            addAbkineticsModel("vax", "IgG", "vax", c("a_vax", "b_vax", "c_vax"), infSerumKinetics),
            addAbkineticsModel("predelta", "IgG", "predelta", c("a_pd", "b_pd", "c_pd"), infSerumKinetics)
        ),
    prior = bind_rows(
        add_par_df("a_vax", -2, 2, "norm",  0, 1), # ab kinetics
        add_par_df("b_vax", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c_vax", 0, 4, "unif", 0,  4), # ab kinetics 
        add_par_df("a_pd", -2, 2, "norm",  0, 1), # ab kinetics
        add_par_df("b_pd", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c_pd", 0, 4, "unif", 0,  4), # ab kinetics
        add_par_df("a_d", -2, 2, "norm",  0, 1), # ab kinetics
        add_par_df("b_d", 0, 1, "norm",  0.3, 0.05), # ab kinetics
        add_par_df("c_d", 0, 4, "unif", 0,  4), # ab kinetics
        add_par_df("wane", 0.0, 0.1, "unif", 0.0, 0.1) # observational model
    )
)

# Define the COP model
copModel <- list( 
        names = c("IgG"),
        model = makeModel(addCopModel("IgG", "delta", c("beta0", "beta1"), copFuncForm,  copLogLikelihood)),
        prior = bind_rows(
            add_par_df("beta0", -10, 10, "unif", -10, 10), # cop model (not used here)
            add_par_df("beta1", -10, 10, "unif", -10, 10)
        ) # cop model (not used here),
)

modeldefinition <- list(
    biomarkers = biomarkers,
    exposureTypes = exposureTypes,
    exposureFitted = exposureFitted,
    observationalModel = observationalModel,
    abkineticsModel = abkineticsModel,
    copModel = copModel,
    exposurePrior = exp_prior_w2,
    exposurePriorType = "empirical"
)

modelW2 <- createSeroJumpModel(gambia_pvnt_w2, gambia_exp_w2, modeldefinition)

````

```{r}

devtools::load_all()

settings <-  list(
    numberChainRuns = 4,
    numberCores = 4,
    iterations = 100,
    burninPosterior = 50,
    thin = 1,
    consoleUpdates = 10,
    onAdaptiveCov = TRUE,
    updatesAdaptiveCov = 10,
    burninAdaptiveCov = 1000,
    covarInitVal = 1e-2, # make very small if struggling to sample to beginning
    covarInitValAdapt = 1e-2, # make very small if struggling to sample to beginning
    covarMaxVal = 1, # decrease if struggling toc sample in the middle
    runParallel = TRUE,
    noGibbsSteps = 1,
    onDebug = FALSE,
    profile = TRUE
)

runRJMCMC(modelW2, settings, "local/transvir/test", "wave2")

postprocessFigs("local/transvir/test", "wave2", 4)

```
# Define settings and run the model, (this won't run as made the data up)

```{r eval = FALSE}

# Define the settings
settings <-  list(
    numberChainRuns = 4,
    numberCores = 4,
    iterations = 200000,
    burninPosterior = 100000,
    thin = 100,
    consoleUpdates = 100,
    onAdaptiveCov = TRUE,
    updatesAdaptiveCov = 100,
    burninAdaptiveCov = 1000,
    updatesAdaptiveTemp = 10,
    covarInitVal = 1e-2, # make very small if struggling to sample to beginning
    covarInitValAdapt = 1e-2, # make very small if struggling to sample to beginning
    covarMaxVal = 1, # decrease if struggling to sample in the middle
    runParallel = TRUE,
    noGibbsSteps = 1
)

# Run the model 
runRJMCMC(modelW2, settings, "test/transvir", "wave2")


```

# Plot all of the posterior plots and calcualte posterior data.frame

```{r}

postprocessFigs_wave2("test/transvir", "wave2", 4)

```